A GPU Sonar Simulator for Automatic Target Recognition

### 2014-05-02 UAM2014/Article/article.tex
la2pdf commit - 2014:05:03	00:23:05	(Sat 3.May)

We have developed a sonar simulator that runs sufficiently fast to create templates adapted to the actual scene as part of the classification process. This is achieved with the aid of the massive computing power in graphics processing units (GPUs) and optimized software libraries for scene rendering. The simulator loads a 3D model of the seafloor and an object class model into OpenGL, where emitted sound waves are modeled with a light source placed at the sonar transmit location. We assume rough, isotropic surfaces reflecting sound energy equally in all directions. This can be modeled with a Lambertian scattering model~\cite{Blake1993,Bell1995}, where the intensity of the backscattered sound only depends on the incidence angle of the transmitted signal onto the model surface. When rendering OpenGL is set up to produce an optical 2D intensity image and depth map that reveals the distance from each pixel to the propagation axis. For maximum flexibility this data is finally combined with OpenCL to produce the image templates.

@@

The next step is to scale, orient and position the objects in the scene, and finally project this scene onto an image plane. This is achieved by applying a set of transformations to each vertex in the model. In OpenGL, the set of transformations is usually formalized like this:
\begin{align*}
\bmat{x\\y\\0\\1}_\text{image} = \P \cdot \V \cdot \M \cdot \bmat{x\\y\\z\\1}_\text{model},
\end{align*}
where is the loaded model, $\bmat{\cdot}_\text{image}$ is the output image, and $\M$, $\V$ and $\P$ are transformation matrices. The first transform we apply is the model matrix $\M$, which scales and rotates the model in its local coordinate system. Then we map this model into world coordinates using the view matrix $\V$. This involves translating the model to its respectful place in the world, and then moving it into the view of the camera. Finally the relevant part of the scene is projected onto the image plane with a projection matrix $\P$.

@@

The image rendered with OpenGL is not in along-track and cross-track coordinates as we want it to be. It does, however, tell us what parts of the scene that are visible from the sonar. In addition, OpenGL can produce a depth map that reveals the distance to each template pixel. This information can be converted to a ranged sonar-like image by simply adding up all the intensity values that share the same depth for each range line. We perform this computation in OpenCL, which allow general purpose programming on GPUs and can interoperate with OpenGL quite nicely. This way we keep all the calculations on the GPU.

@@

Our GPU-based simulator is very fast. On a computer with a six-core Intel Core i7-3930K and a Radeon HD7970 it computes almost 1000 templates per second with a size of 1000x1000 pixels (1 megapixel). This makes it possible to tailor the templates more closely to an image as the parameters needed by the simulator can be directly estimated from the image. In a standard approach, on the other hand, a template library must be created beforehand from a limited set of parameter values. Hence, it may happen that no template in the library is a close fit of the object in the image even though the object actually is a target of interest.

@@

The standard and adaptive templates were then matched to the HISAS image. This is illustrated in Fig. \ref{overlay_cylinder} for a standard template and in Fig. \ref{overlay_cylinder_adaptive} for an adaptive template. These images were created by segmenting the image and the templates into highlight, shadow, and background regions. The regions from a template were then laid on top of the regions of the image to illustrate how well these regions matched. From these figures, it was quite clear the adaptive template obtained a much closer fit to image than the standard template. This was also reflected in the correlation scores (which were created with the method described in \cite{Midelfart2010}) that were 0.613 for the standard template and 0.813 for the adaptive template. Hence, the standard approach was less likely to classify the cylinder in the image correctly as the standard templates were not created specifically for the target.

@@

One way to automatically classify objects in SAS images is to compare the imaged objects with a predefined set of templates. However, this is suboptimal as it is infeasible to create templates for all configurations of seabeds and objects. Instead we have implemented a SAS simulator that creates the templates in delayed real-time based on parameters estimated from the current scene. In our studies this improves the match between the SAS objects and their corresponding template significantly in most cases.

To obtain the delayed real-time performance we implemented the simulator on a GPU. Their theoretical peak performance is typically an order of magnitude higher than a CPU in a comparable price range. This potential is effectively utilized with a well matured and highly optimized graphics processing API such as OpenGL. Our simulator use OpenGL for most of the scene processing.

>>>>

We have developed a synthetic aperture sonar (SAS) simulator that avoids this problem by running sufficiently fast to create templates adapted to the actual scene as part of the classification process. This is achieved with the aid of the massive computing power in graphics processing units (GPUs) and optimized software libraries for scene rendering. The simulator loads a 3D model of the seafloor and an object class model into OpenGL, where emitted sound waves are modeled with a light source placed at the sonar transmit location. We assume rough, isotropic surfaces reflecting sound energy equally in all directions. This can be modeled with a Lambertian scattering model~\cite{Blake1993,Bell1995}, where the intensity of the backscattered sound only depends on the incidence angle of the transmitted signal onto the model surface. When rendering OpenGL is set up to produce an optical 2D intensity image and depth map that reveals the distance from each pixel to the propagation axis. For maximum flexibility this data is finally combined with OpenCL to produce the image templates.

@@

The next step is to scale, orient and position the objects in the scene, and finally project this scene onto an image plane. This is achieved by applying a set of transformations to each vertex in the model. In OpenGL, the set of transformations is usually formalized like this:
\begin{align*}
\bmat{x\\y\\0\\1}_\text{image} = \P \cdot \V \cdot \M \cdot \bmat{x\\y\\z\\1}_\text{model},
\end{align*}
where $\bmat{\cdot}_\text{model}$ is the loaded model, $\bmat{\cdot}_\text{image}$ is the output image, and $\M$, $\V$ and $\P$ are transformation matrices. The first transform we apply is the model matrix $\M$, which scales and rotates the model in its local coordinate system. Then we map this model into world coordinates using the view matrix $\V$. This involves translating the model to its respectful place in the world, and then moving it into the view of the camera. Finally the relevant part of the scene is projected onto the image plane with a projection matrix $\P$.

@@

The ``camera image'' rendered with OpenGL is not in along-track and cross-track coordinates as we want it to be. It does, however, tell us what parts of the scene that are visible from the sonar. OpenGL can also produce a depth map that reveals the distance to each of image pixels. This information can be converted to a ranged sonar-like image by simply adding up all the intensity values that share the same depth for each range line. We perform this computation in OpenCL, which allow general purpose programming on GPUs and can interoperate with OpenGL quite nicely. This way we keep all the calculations on the GPU.

@@

Our GPU-based simulator is very fast. On a computer with a six-core Intel Core i7-3930K and a Radeon HD7970 it computes almost 1000 templates per second, each being 1 megapixel large. This makes it possible to tailor the templates very closely imaged objects. In contrast, in the standard approach a template library must be created beforehand from a limited set of parameter values. Hence, it may happen that no template in the library is a close fit of the object in the image even though the object actually is a target of interest.

@@

The standard and adaptive templates were then matched to the HISAS image. This is illustrated in Fig. \ref{overlay_cylinder} for a standard template and in Fig. \ref{overlay_cylinder_adaptive} for an adaptive template. These images were created by segmenting the image and the templates into highlight, shadow, and background regions. The regions from a template were then laid on top of the regions of the image to illustrate how well these regions matched. Note how the adaptive template obtained a much closer fit to image than the standard template. This was also reflected in the correlation scores (which were created with the method described in \cite{Midelfart2010}) that were 0.613 for the standard template and 0.813 for the adaptive template. Hence, the standard approach was less likely to classify the cylinder in the image correctly as the standard templates were not created specifically for the target.

@@

One way to automatically classify objects in SAS images is to compare the imaged objects with a predefined set of templates. However, this is suboptimal as it is infeasible to create accurate templates for all the relevant configurations of seabeds and objects. To solve this we have implemented a SAS simulator that creates the templates in delayed real-time based on parameters estimated from the current scene. In our studies this improves the match between the SAS objects and their corresponding template significantly in most cases.

To obtain the delayed real-time performance we implemented the simulator on a GPU. These devices have a theoretical peak performance that is typically an order of magnitude higher than CPUs in a comparable price range. We experienced that this potential was effectively utilized with a well matured and highly optimized graphics processing API such as OpenGL. Our simulator use this framework for most of the scene processing.
